---
title: "AMR_cpative_wild"
output: html_document
date: "2026-07-01"
---

```{r setup, include=FALSE}
library(spaMM)
library(Biostrings)
library(MASS)
#library(ape)
library(vegan)
#library(MKinfer)
#library(ecodist)
library(phytools)
#library(nlme)
#library(lme4)
#library(ggrepel)
library(glmmTMB)
#library(patchwork)
#library(ggpubr)
library(data.table)
library(ape)
library(geomorph)
library(RRPP)
library(DHARMa)
library(tidyverse)
library(ggtree)

install.packages("spaMM")

R.version.string

```

#Create base analysis file
```{r}
#Read in MAG metadata
MAG_meta <- read_csv(file = "MAG_metadata_AMR_fixed.csv") %>% rename(genome = MAG_ID)

#Parse AMRfinder outputs
AMR_files <- list.files("amrfinder_out", full.names = TRUE, pattern = "\\.tsv$")

AMR_combined <- rbindlist(
  lapply(
    AMR_files, 
    function(f){fread(f)[, genome := str_remove(basename(f), "\\.tsv$")]}
         ), 
  fill = TRUE
)

#AMR_combined %>% write_csv(file = "amr_out_combined.csv")

#read in dRep clustering data
drep_clustering <- read_csv(file = "MAGs/pass/drep_out/data_tables/Cdb.csv") %>% mutate(genome = str_remove(genome, "\\.fa$"))


#Combine amr outs, mag meta, and drep clustering
combined_amr_data <- left_join(AMR_combined, select(MAG_meta, genome, common_host, status, completeness, contamination, phylum, class, family, genus, species, sample_id, study, run_thru_sn_mg_pipeline, country, city), by = "genome") %>% left_join(., select(drep_clustering, genome, secondary_cluster), by = "genome") %>% select(genome,common_host, status, secondary_cluster,everything())


#also need drep cluster membership
cluster_members <- left_join(select(drep_clustering, genome, secondary_cluster), select(MAG_meta, genome, common_host, status))


```

#Wildlife MAG tree annotation (-humans and -Koala)
```{r}
#Full MAG tree could be visualized, but not really with the way we want to annotate it. Tree has ~6000 tips, so any information like AMR abundance would be a 1/6000th slice of the tree. Going to split tree into per host trees. This makes the largest tree ~2000 tips, still a bit hard to visualize annotations, but could either show off Hosts with fewer MAGs or zoom into interesting parts of the tree

#number of MAGs in full tree per common host
AMR_combined %>% 
  filter(Scope == "core" & Type == "AMR") %>% 
  group_by(genome) %>% 
  summarise(AMR_abundance = n()) %>% 
  full_join(., select(MAG_meta, genome, common_host, status), by = "genome") %>% 
  mutate(AMR_abundance = replace_na(AMR_abundance, 0)) %>% 
  filter(common_host != "Human" & common_host != "Koala") %>% 
  filter(genome %in% full_tree$tip.label) %>% 
  group_by(common_host) %>% 
  summarise(n = n())



#definitely easier ways to make a list of hosts, but this makes sure each host is actually in the tree. 
host_list <- AMR_combined %>% 
  filter(Scope == "core" & Type == "AMR") %>% 
  group_by(genome) %>% 
  summarise(AMR_abundance = n()) %>% 
  full_join(., select(MAG_meta, genome, common_host, status), by = "genome") %>% 
  mutate(AMR_abundance = replace_na(AMR_abundance, 0)) %>% 
  filter(common_host != "Human" & common_host != "Koala") %>% 
  filter(genome %in% full_tree$tip.label) %>% 
  distinct(common_host) %>% 
  pull()

per_host_tree_list <- list()
per_host_tree_annotate_list <- list()

for (host in host_list) {
  
  tmp_host_tree_df <- AMR_combined %>% 
    filter(Scope == "core" & Type == "AMR") %>% 
    group_by(genome) %>% 
    summarise(AMR_abundance = n()) %>% 
    full_join(., select(MAG_meta, genome, common_host, status), by = "genome") %>% 
    mutate(AMR_abundance = replace_na(AMR_abundance, 0)) %>% 
    filter(common_host != "Human" & common_host != "Koala") %>% 
    filter(genome %in% full_tree$tip.label) %>% 
    left_join(., select(MAG_meta, genome, phylum, class, family, genus, species)) %>% 
    filter(common_host == host)
  
  
  tmp_tree <- keep.tip.phylo(full_tree, tmp_host_tree_df$genome)
  
  per_host_tree_list[[host]] <- tmp_tree
  
  per_host_tree_annotate_list[[host]] <- select(tmp_host_tree_df, genome, AMR_abundance, common_host, status, phylum)
  
}

per_host_tree_list[[host]]
per_host_tree_annotate_list[[host]]


#look into these with ggtree
```

#MAG PGLS
```{r}

#MAG tree with ALL the samples from the study (except is missing some). Might want to re-infer tree with more specific set of MAGs
full_tree <- read.tree(file = "all_animal+drep95_human.nwk")

#Full AMR abundance including MAGs with 0 abundance. ~70 MAGs removed because they are not in the tree. Ones not in tree all had 0 AMR abundance and were from a variety of host species and studies. At least a couple samples from which missing MAGs were derived did include other MAGs in tree, so it probably wasn't a sample issue.
phylo_anova_df <- AMR_combined %>% 
  filter(Scope == "core" & Type == "AMR") %>% 
  group_by(genome) %>% 
  summarise(AMR_abundance = n()) %>% 
  full_join(., select(MAG_meta, genome, common_host, status), by = "genome") %>% 
  mutate(AMR_abundance = replace_na(AMR_abundance, 0)) %>% 
  filter(common_host != "Human" & common_host != "Koala") %>% 
  filter(genome %in% full_tree$tip.label) %>% 
  column_to_rownames(var = "genome")

#Prune full MAG tree down to those included in the current analysis (Core AMR genes from Animal MAGs - Koala).
phylo_anova_pruned_tree <- keep.tip.phylo(full_tree, rownames(phylo_anova_df))



phylo_anova_df <- phylo_anova_df[phylo_anova_pruned_tree$tip.label,]
pgls_df <- phylo_anova_df %>% rownames_to_column(var = "genome") %>% mutate(AMR_present = AMR_abundance > 0)

AMR_binary <- pgls_df$AMR_present
names(AMR_binary) <- pgls_df$genome

AMR <- pgls_df$AMR_abundance
names(AMR) <- pgls_df$genome

status <- pgls_df$status
names(status) <- pgls_df$genome

common_host <- pgls_df$common_host
names(common_host) <- pgls_df$genome


pgls_df_geomorph <- geomorph.data.frame(
    AMR_abundance = AMR,
    status = status,
    common_host = common_host,
    AMR_binary = AMR_binary,
    phy = phylo_anova_pruned_tree
)

pgls_model <- procD.pgls(
    AMR ~ status + common_host,
    phy = phy,
    data = pgls_df_geomorph,
    iter = 9999
)

lm_model <- procD.lm(
    AMR ~ status + common_host,
    data = pgls_df_geomorph,
    iter = 9999
)


pgls_model_summary <- summary(pgls_model)

lm_model_summary <- summary(lm_model)



# This version of the MAG phylogenetic tree doesn't seem to be appropriate for some phylogenetic comparative methods. It looks like the abundance of highly similar MAGs at least in terms of branch length are causing problems. For now there is a mixed model that does not take into account the MAG phylogeny. 

# Phylogenetic covariance/correlation matrix for pglmm
#tree_corr <- vcv(phylo_anova_pruned_tree, corr = TRUE)

#pglmm_df <- pgls_df


#pglmm_df$genome <- factor(pglmm_df$genome, levels = rownames(tree_corr))

#pglmm_nb_model <- fitme(AMR_abundance ~ status + (1 | common_host) + corrMatrix(1 | genome), data = pglmm_df, family = negbin(), corrMatrix = tree_corr)

#summary(pglmm_nb_model)

glmm_model_summary <- glmmTMB(AMR_abundance ~ status + (1 | common_host), family = nbinom2,data = pgls_df) %>% summary()
```



#Full sample analysis below
#read in AMR data and Metadata
```{r}
contig_info <- read_table(file = "contigs/contig_stats.txt") %>% 
  mutate(sample = str_replace(file, ".contigs.fasta", ""), len_mb = sum_len/1000000) %>% 
  select(-file,-format,-type)

sample_meta <- read_csv(file = "sample_metadata.csv") %>% 
  left_join(.,contig_info, by = "sample") %>% 
  filter(common_host != "Koala" & !is.na(common_host))


AMR_files_contigs <- list.files("amrfinder_contigs_out", full.names = TRUE, pattern = "\\.tsv$")

AMR_combined_contigs <- rbindlist(
  lapply(
    AMR_files_contigs, 
    function(f){fread(f)[, sample := str_remove(basename(f), "\\.contigs.tsv$")]}
         ), 
  fill = TRUE
) %>% select(sample, everything())


sample_AMR <- full_join(AMR_combined_contigs, sample_meta, by = "sample") %>% 
  filter(!is.na(common_host) & common_host != "Koala") %>% 
  select(sample, "Element symbol", Scope, Type, common_host, status, City, len_mb) %>% 
  rename(gene_id = "Element symbol")


```


#Bray distance PERMANOVA on full sample set
```{r}
bray_df_raw <- sample_AMR %>% 
  filter(Scope == "core" & Type == "AMR") %>%
  count(sample, gene_id) %>% 
  pivot_wider(names_from = gene_id, values_from = n, values_fill = 0) %>%
  column_to_rownames(var = "sample")

bray_dist_raw <- vegdist(bray_df_raw, method = "bray")
#Don't need this here, but can use for pcoa later if needed.
bray_pcoa_raw <- pcoa(bray_dist_raw)

anova_df_raw <- sample_meta %>% filter(sample %in% rownames(bray_df_raw))

bray_df_norm <- sample_AMR %>% 
  filter(Scope == "core" & Type == "AMR") %>%
  count(sample, gene_id) %>% 
  pivot_wider(names_from = gene_id, values_from = n, values_fill = 0) %>% 
  left_join(., select(sample_meta, sample, len_mb)) %>% 
  mutate(across(-c(sample, len_mb), ~.x/len_mb)) %>% 
  select(-len_mb) %>% 
  column_to_rownames(var = "sample")

bray_dist_norm <- vegdist(bray_df_norm, method = "bray")

#Don't need this here, but can use for pcoa later if needed.
bray_pcoa_norm <- pcoa(bray_dist_norm)

anova_df_norm <- sample_meta %>% filter(sample %in% rownames(bray_df_norm))



raw_abundance_Bray_permnova <- adonis2(bray_dist_raw ~ Country + status + common_host, data = anova_df_raw, permutations = 9999, strata = anova_df_raw$common_host, by = "margin")

norm_abundance_Bray_permnova <- adonis2(bray_dist_norm ~ Country + status + common_host, data = anova_df_norm, permutations = 9999, strata = anova_df_norm$common_host, by = "margin")



raw_abundance_Bray_permnova
norm_abundance_Bray_permnova

```


#Abundance GLMM on full sample set
```{r}
total_abundance_sample_df <- bray_df_raw %>% 
  mutate(abundance = rowSums(across(everything()))) %>% 
  select(abundance) %>% 
  rownames_to_column(var = "sample") %>% 
  full_join(., select(sample_meta, sample, common_host, status, Country, len_mb), by = "sample") %>% 
  mutate(abundance = coalesce(abundance, 0), norm_abundance = abundance/len_mb)


total_abundance_sample_glmm <- glmmTMB(abundance ~ status + offset(log(len_mb)) + (1 | common_host) + (1 | Country), family = nbinom2, data = total_abundance_sample_df) %>% summary()

total_abundance_sample_glmm


```


#Bray distance PERMANOVA on per host species 
```{r}

#Many of the hosts will fail in this analysis due to low sample size. Only samples with at least 1 AMR gene are included here, so many hosts have their sample sizes reduced further. If a host species ends up with 1 or 0 samples of either wild or captive status the sample PERMANOVA will fail.
taxa_list <- sample_AMR %>% distinct(common_host) %>% pull(common_host)
adonis_results_raw_list <- list()
adonis_results_norm_list <- list()


for (level in taxa_list) {


  bray_df_raw <- sample_AMR %>% 
    filter(Scope == "core" & Type == "AMR") %>%
    filter(common_host == level) %>% 
    count(sample, gene_id) %>% 
    pivot_wider(names_from = gene_id, values_from = n, values_fill = 0) %>%
    column_to_rownames(var = "sample")

  bray_dist_raw <- vegdist(bray_df_raw, method = "bray")
  

  anova_df_raw <- sample_meta %>% filter(sample %in% rownames(bray_df_raw))

  bray_df_norm <- sample_AMR %>% 
    filter(Scope == "core" & Type == "AMR") %>%
    filter(common_host == level) %>%
    count(sample, gene_id) %>% 
    pivot_wider(names_from = gene_id, values_from = n, values_fill = 0) %>% 
    left_join(., select(sample_meta, sample, len_mb)) %>% 
    mutate(across(-c(sample, len_mb), ~.x/len_mb)) %>% 
    select(-len_mb) %>% 
    column_to_rownames(var = "sample")

  bray_dist_norm <- vegdist(bray_df_norm, method = "bray")
  
  anova_df_norm <- sample_meta %>% filter(sample %in% rownames(bray_df_norm))
  
  if (n_distinct(anova_df_raw$Country) > 1) {
    model_formula_raw <- bray_dist_raw ~ Country + status
  } else {
    model_formula_raw <- bray_dist_raw ~ status
  }
  
  if (n_distinct(anova_df_norm$Country) > 1) {
    model_formula_norm <- bray_dist_norm ~ Country + status
  } else {
    model_formula_norm <- bray_dist_norm ~ status
  }
  
  set.seed(123)
  
  adonis_results_raw_list[[level]] <- adonis2(model_formula_raw, data = anova_df_raw, permutations = 9999, by = "margin")
  
  set.seed(123)
  
  adonis_results_norm_list[[level]] <- adonis2(model_formula_norm, data = anova_df_norm, permutations = 9999, by = "margin")
  

}


adonis_results_raw_list
adonis_results_norm_list

```


#general plotting - without zeros
```{r}
AMR_summary <- sample_AMR %>% 
  group_by(common_host) %>% 
  summarize(
    n_wild = n_distinct(sample[status == "wild"]), 
    n_captive = n_distinct(sample[status == "captive"]), 
    wild_AMR = sum(status == "wild"),  
    captive_AMR = sum(status == "captive"), 
    wild_AMR_per_ind = wild_AMR/n_wild, 
    captive_AMR_per_ind = captive_AMR/n_captive)

sample_AMR %>% 
  filter(Scope == "core" & Type == "AMR") %>%
  count(sample, status) %>%
  ggplot(aes(x = status, y = n)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1)

sample_AMR

sample_AMR_plot <- sample_AMR %>%
  filter(Scope == "core" & Type == "AMR") %>%
  count(sample, common_host, status, len_mb, name = "AMR_abundance") %>%
  mutate(AMR_abundance_norm = AMR_abundance / len_mb)

sample_AMR_plot


# 1. Raw abundance: all captive vs all wild
raw_all_boxplot <- ggplot(sample_AMR_plot, aes(x = status, y = AMR_abundance)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) +
  labs(x = NULL, y = "AMR abundance") +
  theme_classic()

# 2. Raw abundance: captive vs wild within each host
raw_host_species_plot <- ggplot(sample_AMR_plot, aes(x = common_host, y = AMR_abundance, fill = status)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  labs(x = NULL, y = "AMR abundance") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 3. Normalized abundance: all captive vs all wild
norm_all_boxplot <- ggplot(sample_AMR_plot,aes(x = status, y = AMR_abundance_norm)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) +
  labs(x = NULL, y = "AMR abundance per Mb") + 
  theme_classic()

# 4. Normalized abundance: captive vs wild within each host
norm_host_species_plot <- ggplot(sample_AMR_plot,aes(x = common_host, y = AMR_abundance_norm, fill = status)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  labs(x = NULL, y = "AMR abundance per Mb") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


raw_all_boxplot
raw_host_species_plot
norm_all_boxplot
norm_host_species_plot
```


#general plotting with zeros - do this per MAG as well
```{r}
sample_AMR_plot_zeros <- sample_meta %>%
  select(sample, common_host, status, len_mb) %>%
  left_join(sample_AMR %>%filter(Scope == "core" & Type == "AMR") %>% 
  count(sample, name = "AMR_abundance"), by = "sample") %>%
  mutate(AMR_abundance = replace_na(AMR_abundance, 0),AMR_abundance_norm = AMR_abundance / len_mb) %>% 
  filter(common_host != "Koala" & !is.na(common_host))


# 1. Raw abundance: all captive vs all wild
raw_all_boxplot_zeros <- ggplot(sample_AMR_plot_zeros, aes(x = status, y = AMR_abundance)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) +
  labs(x = NULL, y = "AMR abundance") +
  theme_classic()

# 2. Raw abundance: captive vs wild within each host
raw_host_species_plot_zeros <- ggplot(sample_AMR_plot_zeros, aes(x = common_host, y = AMR_abundance, fill = status)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  labs(x = NULL, y = "AMR abundance") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 3. Normalized abundance: all captive vs all wild
norm_all_boxplot_zeros <- ggplot(sample_AMR_plot_zeros,aes(x = status, y = AMR_abundance_norm)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) +
  labs(x = NULL, y = "AMR abundance per Mb") + 
  theme_classic()

# 4. Normalized abundance: captive vs wild within each host
norm_host_species_plot_zeros <- ggplot(sample_AMR_plot_zeros,aes(x = common_host, y = AMR_abundance_norm, fill = status)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  labs(x = NULL, y = "AMR abundance per Mb") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

raw_all_boxplot_zeros
raw_host_species_plot_zeros

norm_all_boxplot_zeros
norm_host_species_plot_zeros

```


#PCOA ploting
```{r}
var_explained <- bray_pcoa$values$Relative_eig * 100
var_explained
coords <- as.data.frame(bray_pcoa$vectors[,1:2])
coords$sample <- rownames(coords)

coords <- left_join(coords, sample_meta, by = "sample")
coords

ggplot(coords, aes(x = Axis.1,y = Axis.2, color = status)) +
  geom_point(size = 3) +
  xlab(paste0("PCoA1 (", round(var_explained[1],1), "%)")) +
  ylab(paste0("PCoA2 (", round(var_explained[2],1), "%)")) +
  theme_classic()

```



#Full sample AMR class PERMANOVA
```{r}

AMR_class_df <- full_join(AMR_combined_contigs, sample_meta, by = "sample") %>%
  filter(!is.na(common_host) & common_host != "Koala") %>% 
  filter(Scope == "core" & Type == "AMR") %>% 
  separate_longer_delim(Class, delim = "/") %>% 
  count(sample, Class, name = "n_hits") %>% 
  pivot_wider(names_from = "Class", values_from = "n_hits", values_fill = 0) %>% 
  column_to_rownames(var = "sample")

ARM_class_meta <- sample_meta %>% filter(sample %in% rownames(AMR_class_df))

ARM_class_bray <- vegdist(AMR_class_df, method = "bray")


AMR_class_abundance_Bray_permnova <- adonis2(ARM_class_bray ~ Country + status + common_host, data = ARM_class_meta, permutations = 9999, strata = ARM_class_meta$common_host, by = "margin")

AMR_class_abundance_Bray_permnova


```

#per class GLMM
```{r}
# Convert antibiotic class counts from wide to long format
AMR_class_long_df <- AMR_class_df %>% rownames_to_column(var = "sample") %>% 
  pivot_longer(cols = -sample,names_to = "antibiotic_class", values_to = "n") %>%
  mutate(n = coalesce(n, 0)) %>%left_join(select(sample_meta, sample, len_mb, common_host, status, Country),by = "sample") %>%
  filter(common_host != "Koala" & !is.na(common_host))


# Classes present in at least 15 samples
AMR_class_list <- AMR_class_long_df %>%
  filter(n > 0) %>%
  distinct(sample, antibiotic_class) %>%
  count(antibiotic_class, name = "prevalence") %>%
  filter(prevalence >= 15) %>%
  pull(antibiotic_class)


# Containers for models/results
AMR_class_glmm_poisson_model <- list()
AMR_class_glmm_nbinom_model <- list()

AMR_class_glmm_poisson_results <- list()
AMR_class_glmm_nbinom_results <- list()

AMR_class_glmm_selected_model_results <- data.frame()


for (drug_class in AMR_class_list) {

  AMR_class_GLMM_df <- AMR_class_long_df %>% filter(antibiotic_class == drug_class)

  
  # Poisson model
  AMR_class_glmm_poisson_model[[drug_class]] <- glmmTMB(n ~ status + offset(log(len_mb)) + (1 | common_host) + (1 | Country), family = poisson, data = AMR_class_GLMM_df)

  
  # Negative binomial model
  AMR_class_glmm_nbinom_model[[drug_class]] <- glmmTMB(n ~ status +offset(log(len_mb)) + (1 | common_host) + (1 | Country), family = nbinom2, data = AMR_class_GLMM_df)

  
  AMR_class_glmm_poisson_results[[drug_class]] <- summary(AMR_class_glmm_poisson_model[[drug_class]])

  AMR_class_glmm_nbinom_results[[drug_class]] <-summary(AMR_class_glmm_nbinom_model[[drug_class]])

  
  poisson_aic <- AMR_class_glmm_poisson_results[[drug_class]]$AICtab[[1]]

  negbinom_aic <- AMR_class_glmm_nbinom_results[[drug_class]]$AICtab[[1]]

  
  # Select model using AIC
  if (is.na(negbinom_aic) || poisson_aic <= negbinom_aic) {

    selected_model <- "poisson"

    selected_results <- AMR_class_glmm_poisson_results[[drug_class]]

    model_qc <- simulateResiduals(AMR_class_glmm_poisson_model[[drug_class]])

  } 
  
  else {

    selected_model <- "negative_binomial"

    selected_results <- AMR_class_glmm_nbinom_results[[drug_class]]

    model_qc <- simulateResiduals(AMR_class_glmm_nbinom_model[[drug_class]])
  }

  
  # DHARMa diagnostics
  Uniformity_pvalue <- testUniformity(model_qc, plot = FALSE)$p.value

  Dispersion_pvalue <- testDispersion(model_qc, plot = FALSE)$p.value

  zero_inflation_pvalue <- testZeroInflation(model_qc, plot = FALSE)$p.value

  
  # Wild vs captive coefficient
  estimate <-selected_results$coefficients$cond["statuswild", "Estimate"]

  p_value <- selected_results$coefficients$cond["statuswild", "Pr(>|z|)"]

  
  # Store results
  AMR_class_glmm_selected_model_results <-
    bind_rows(AMR_class_glmm_selected_model_results, data.frame(
        antibiotic_class = drug_class,
        estimate_wild = estimate,
        fold_change = exp(estimate),
        p_value = p_value,
        model_qc_flag = Uniformity_pvalue < 0.05 | Dispersion_pvalue < 0.05 | zero_inflation_pvalue < 0.05,
        selected_model = selected_model,
        poisson_aic = poisson_aic,
        negbinom_aic = negbinom_aic,
        count_mean = mean(AMR_class_GLMM_df$n),
        count_var = var(AMR_class_GLMM_df$n),
        Uniformity_pvalue = Uniformity_pvalue,
        Dispersion_pvalue = Dispersion_pvalue,
        zero_inflation_pvalue = zero_inflation_pvalue
      )
    )
}


# Benjamini-Hochberg multiple-testing correction
AMR_class_glmm_selected_model_results$fdr <- p.adjust(AMR_class_glmm_selected_model_results$p_value, method = "BH")


AMR_class_glmm_selected_model_results <-AMR_class_glmm_selected_model_results %>% select(antibiotic_class, estimate_wild, fold_change, p_value, fdr, everything())


AMR_class_glmm_selected_model_results

AMR_class_glmm_selected_model_results %>% filter(fdr <= 0.05)


```

#Per gene GLMM
```{r}
#GLMM df base
AMR_trees_base_df <- left_join(AMR_combined_contigs, sample_meta, by = "sample") %>% 
  filter(common_host != "Koala" & !is.na(common_host)) %>%
  filter(Scope == "core" & Type == "AMR") %>% 
  rename(gene_id = "Element symbol")

#GLMM per gene
#Gene has to be present in at least 15 samples to be included.
AMR_genes_list <- AMR_trees_base_df %>%
  distinct(sample, gene_id) %>%
  count(gene_id, name = "prevalence") %>%
  filter(prevalence >= 15) %>%
  pull(gene_id)

AMR_genes_glmm_poisson_model <- list()
AMR_genes_glmm_nbinom_model <- list() 
AMR_genes_glmm_poisson_results <- list()
AMR_genes_glmm_nbinom_results <- list()
AMR_genes_glmm_selected_model_results <- data.frame()


for (gene in AMR_genes_list){

  AMR_trees_GLMM_df <- AMR_trees_base_df %>% 
    select(sample, gene_id, common_host, status, len_mb) %>% 
    filter(gene_id == gene) %>% 
    count(sample, gene_id) %>% 
    full_join(., select(sample_meta, sample, len_mb, common_host, status, Country)) %>% 
    mutate(n = coalesce(n, 0))
  
  
  #Create glmm models using both poisson and negative binomial distributions
  AMR_genes_glmm_poisson_model[[gene]] <- glmmTMB(n ~ status + offset(log(len_mb)) + (1 | common_host) + (1 | Country), family = poisson, data = AMR_trees_GLMM_df)
  
  AMR_genes_glmm_nbinom_model[[gene]] <- glmmTMB(n ~ status + offset(log(len_mb)) + (1 | common_host) + (1 | Country), family = nbinom2, data = AMR_trees_GLMM_df)
  
  AMR_genes_glmm_poisson_results[[gene]] <- summary(AMR_genes_glmm_poisson_model[[gene]])
  
  AMR_genes_glmm_nbinom_results[[gene]] <- summary(AMR_genes_glmm_nbinom_model[[gene]])
  
  poisson_aic <- AMR_genes_glmm_poisson_results[[gene]]$AICtab[[1]]
  negbinom_aic <- AMR_genes_glmm_nbinom_results[[gene]]$AICtab[[1]]

  
  #select appropriate model using AIC and then run some qc checks
  if (is.na(negbinom_aic) || poisson_aic <= negbinom_aic) {
  
      selected_model <- "poisson"
      selected_results <- AMR_genes_glmm_poisson_results[[gene]]
      
      
      model_qc <- simulateResiduals(AMR_genes_glmm_poisson_model[[gene]])
      Uniformity_pvalue <- testUniformity(model_qc, plot = FALSE)$p.value
      Dispersion_pvalue <- testDispersion(model_qc, plot = FALSE)$p.value
      zero_infliation_pvalue <- testZeroInflation(model_qc, plot = FALSE)$p.value
  
    } else {
  
      selected_model <- "negative_binomial"
      selected_results <- AMR_genes_glmm_nbinom_results[[gene]]
      
      model_qc <- simulateResiduals(AMR_genes_glmm_nbinom_model[[gene]])
      Uniformity_pvalue <- testUniformity(model_qc, plot = FALSE)$p.value
      Dispersion_pvalue <- testDispersion(model_qc, plot = FALSE)$p.value
      zero_infliation_pvalue <- testZeroInflation(model_qc, plot = FALSE)$p.value
  
    }

  estimate <- selected_results$coefficients$cond["statuswild", "Estimate"]

  p_value <- selected_results$coefficients$cond["statuswild", "Pr(>|z|)"]

  #create df of sample information, model results/information, and QC
  AMR_genes_glmm_selected_model_results <- bind_rows(AMR_genes_glmm_selected_model_results, data.frame(
    gene_id = gene,
    estimate_wild = estimate,
    fold_change = exp(estimate),
    p_value = p_value,
    model_qc_flag = Uniformity_pvalue < 0.05 | Dispersion_pvalue < 0.05 | zero_infliation_pvalue < 0.05,
    selected_model = selected_model,
    poisson_aic = poisson_aic,
    negbinom_aic = negbinom_aic,
    count_mean = mean(AMR_trees_GLMM_df$n),
    count_var = var(AMR_trees_GLMM_df$n),
    Uniformity_pvalue = Uniformity_pvalue,
    Dispersion_pvalue = Dispersion_pvalue,
    zero_infliation_pvalue = zero_infliation_pvalue
    
  ))
  
  }

#Multiple comparisons adjustment to p_values using Benjamini–Hochberg correction
AMR_genes_glmm_selected_model_results$fdr <- p.adjust(AMR_genes_glmm_selected_model_results$p_value, method = "BH")

AMR_genes_glmm_selected_model_results <- AMR_genes_glmm_selected_model_results %>% 
  select(gene_id, estimate_wild, fold_change, p_value, fdr, everything())

#Per AMR gene wild/captive association
AMR_genes_glmm_selected_model_results %>% filter(fdr <= .05)

#little histogram to visualize gene counts across samples. Will display counts for gene that is currently described by the gene var in the above loop. 
#AMR_trees_GLMM_df %>% ggplot(aes(x = n)) +
#  geom_histogram()

```

#Human plus animal AMR gene trees
```{r}

sample_meta_humans <- read_csv(file = "sample_metadata.csv") %>% 
  left_join(.,contig_info, by = "sample") %>% 
  mutate(common_host = replace_na(common_host, "human"), status = replace_na(status, "human")) %>% 
  filter(common_host != "Koala")

#slighty different from similar dataframes in that it includes humans and does not include samples without any AMR (uses left_join rather than full_join). 
sample_AMR_humans <- left_join(AMR_combined_contigs, sample_meta_humans, by = "sample") %>% 
  filter(common_host != "Koala") %>% 
  select(sample, "Element symbol", Scope, Type, common_host, status, City, len_mb, "Protein id", Start, Stop, Strand, "Contig id", "% Coverage of reference") %>% 
  rename(contig_id = "Contig id", gene_id = "Element symbol", coverage_of_reference = "% Coverage of reference", protein_id = "Protein id") %>% 
  filter(Scope == "core" & Type == "AMR")

#MAG df used as a key to add bacterial taxonomy of those genes that were included in a MAG
MAG_key_df <- combined_amr_data %>% 
  filter(Scope == "core" & Type == "AMR") %>% 
  rename(contig_id = "Contig id", sample = sample_id, gene_id = "Element symbol", coverage_of_reference = "% Coverage of reference") %>% 
  select(sample, gene_id, Start, Stop, Strand, coverage_of_reference, genome, phylum, class, family, genus, species)

#Add bacterial taxonomic info to genes that were included in a MAG. Genes without taxonomy info are on contigs that were not included in a MAG.
sample_AMR_humans <- left_join(sample_AMR_humans, MAG_key_df, by = c("sample", "gene_id", "Start", "Stop", "Strand", "coverage_of_reference"))

#I want to know for each gene how many human, wild, captive counts.
sample_gene_summary <- sample_AMR_humans %>% 
  group_by(gene_id) %>% 
  summarise(n = n(), n_wild = sum(status == "wild"), n_captive = sum(status == "captive"), n_human = sum(status == "human"))


sample_gene_tree_list <- sample_gene_summary %>% 
  filter(n_wild >= 5 & n_captive >= 5 & n_human >= 5) %>% 
  pull(gene_id)



sample_hits <- sample_AMR_humans %>% filter(gene_id %in% sample_gene_tree_list)
AMR_seqs_df <- list()


for (samp in unique(sample_hits$sample)){
  
  seqs_df_tmp <- sample_hits %>% filter(sample == samp)
  
  fna_file <- paste0("contig_annotations/", samp, ".contigs/", samp, ".contigs.fna")
  
  fna_seqs <- readDNAStringSet(fna_file)
  
  names(fna_seqs) <- sub(" .*", "", names(fna_seqs))
  
  
  
  AMR_seqs_df[[samp]] <-seqs_df_tmp %>% 
    mutate(sequence = pmap_chr(pick(contig_id, Start, Stop, Strand), 
    function(contig_id, Start, Stop, Strand) {
      
      seq <- subseq(fna_seqs[[contig_id]], start = Start, end = Stop)
                        
      if (Strand == "-") {
        seq <- reverseComplement(seq)
        }
      
      as.character(seq)
  }))
}


AMR_seqs_df <- bind_rows(AMR_seqs_df)

AMR_seqs_df

gene_lookup <- AMR_seqs_df %>%
  distinct(gene_id) %>%
  mutate(gene_num = paste0("gene_", row_number()))

AMR_seqs_df <- AMR_seqs_df %>%
  left_join(gene_lookup, by = "gene_id") %>%
  group_by(gene_id) %>%
  mutate(seq_id = paste0("seq_", row_number())) %>%
  ungroup()

AMR_seqs_df %>% 
  split(.$gene_num) %>% iwalk(function(df, gene) {

    seqs <- DNAStringSet(df$sequence)

    names(seqs) <- df$seq_id

    writeXStringSet(
      seqs,
      paste0("AMR_gene_fastas/", gene, ".fasta")
    )
  })

#AMR_seqs_df %>% select(sample, gene_id, gene_num, seq_id, common_host, status, City, genome, phylum, class, family, genus, species) %>% write_delim(file = "AMR_gene_fastas/lookup_table.txt")

#Command used to align fasta files
#parallel -j 5 '/programs/mafft/bin/mafft --thread 5 --localpair --maxiterate 1000 {} > {/.}.align' ::: *.fasta

#Command used to infer trees from alignments
#parallel -j 5 /programs/iqtree-2.4.0/bin/iqtree2 -s {} -m MFP --seed 0 --prefix {/.} -T 10 -B 1000 ::: *.align
```


#Gene trees vis
```{r}
AMR_seqs_df %>% select(sample, gene_id, gene_num, seq_id, common_host, status, City, genome, phylum, class, family, genus, species)


AMR_seqs_df
make_gene_tree <- function(tree){

gene_tree<- read.tree(paste0("AMR_gene_fastas/iqtree_outs/", tree, ".treefile"))

gene_meta <- AMR_seqs_df %>%
  filter(gene_num == tree) %>%
  select(seq_id, sample, status, common_host) %>% 
  rename(label = seq_id)

ggtree(gene_tree) %<+% gene_meta + geom_tippoint(aes(color = status),size = 1)

}



make_gene_tree("gene_25")

```


